Scagnostics and Projection Pursuit

### 2018-01-31 Draft/Draft.Rmd
Small additions in the text.

For the following illustrations we work with the particle physics dataset (ATLAS Collaboration, 2015).
This data includes a number of predictions for observables (in agreement with current limits) and the so-called finetuning measure for a large set of parameter points in a supersymmetric model.
The observables describe predictions for decay branching ratios and other precision observables that are generally used to constrain new models of particle physics.
We use the log transformed values for the variables XYZ but not the finetuning, and we rescale all variables to a [0,1] interval
to obtain comparable scales.
Contributions to the variable gminus2 can be negative or positive, we therefore add a small shift before the log transformation of this variable.

@@

## Additional indices to consider
FIXME load Katrins package and try the index functions she implemented

>>>>

For the following illustrations we work with the particle physics dataset (ATLAS Collaboration, 2015).
This data includes a number of predictions for observables (in agreement with current limits) and the so-called finetuning measure for a large set of parameter points in a supersymmetric model.
The observables describe predictions for decay branching ratios and other precision observables that are generally used to constrain new models of particle physics.
We use the log transformed values for the variables XYZ but not the finetuning, and we rescale all variables to a [0,1] interval
to obtain comparable scales.
Contributions to the variable gminus2 can be negative or positive, we therefore add a small shift before the log transformation of this variable.
Maybe include GGally pairs plot here?

@@

## Additional indices to consider

Additions to the standard scagnostics measures have been implemented in the package mbgraphics, based on splines (spline2d) and distance correlation (dcor2d).
Give definitions here..
We repeat the study from before, we first test in Figure ADDREF the behaviour of the index values over 100 interpolated projections that are randomly selected via the grand tour algorithm (the same projections as considered above).
Note that here the measures are calculated on non-binned data.
The general behaviour seems similar to the scagnostics monotonic index.
The observed behaviour is smooth (much in contrast to the behaviour of the scagnostics measures) and we next test the behaviour in a guided tour, comparing again also the first and final projection. (which index to select, will try both below...)
It seems they both maximise correlation, this might not be too useful?
```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
```
```{r guided-spline2d, include=FALSE}
```
```{r guided-spline2d-plot, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
```
```{r guided-dcor2d, include=FALSE}
```
```{r guided-dcor2d-plot, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
```

Physics: what is final projection, what is special about outlying points?

maybe move timing to the end, add also these indices in the same plot (instead of testing several samples, behaviour anyway similar)

### 2018-01-31 Draft/Draft.tex
Small additions in the text.

For the following illustrations we work with the particle physics
dataset (ATLAS Collaboration, 2015). This data includes a number of
predictions for observables (in agreement with current limits) and the
so-called finetuning measure for a large set of parameter points in a
supersymmetric model. The observables describe predictions for decay
branching ratios and other precision observables that are generally used
to constrain new models of particle physics. We use the log transformed
values for the variables XYZ but not the finetuning, and we rescale all
variables to a {[}0,1{]} interval to obtain comparable scales.
Contributions to the variable gminus2 can be negative or positive, we
therefore add a small shift before the log transformation of this
variable.

@@

FIXME load Katrins package and try the index functions she implemented

Additions to the standard scagnostics measures have been implemented in
the package mbgraphics, based on splines (spline2d) and distance
correlation (dcor2d). Give definitions here.. We repeat the study from
before, we first test in Figure ADDREF the behaviour of the index values
over 100 interpolated projections that are randomly selected via the
grand tour algorithm. Note that here the measures are calculated on
non-binned data. The observed behaviour is smooth (much in contrast to
the behaviour of the scagnostics measures) and we next test the
behaviour in a guided tour, comparing again also the first and final
projection. (which index to select, will try both below\ldots{}) It
seems they both maximise correlation, this might not be too useful
here\ldots{}

@@

>>>>

For the following illustrations we work with the particle physics
dataset (ATLAS Collaboration, 2015). This data includes a number of
predictions for observables (in agreement with current limits) and the
so-called finetuning measure for a large set of parameter points in a
supersymmetric model. The observables describe predictions for decay
branching ratios and other precision observables that are generally used
to constrain new models of particle physics. We use the log transformed
values for the variables XYZ but not the finetuning, and we rescale all
variables to a {[}0,1{]} interval to obtain comparable scales.
Contributions to the variable gminus2 can be negative or positive, we
therefore add a small shift before the log transformation of this
variable. Maybe include GGally pairs plot here?

@@

Additions to the standard scagnostics measures have been implemented in
the package mbgraphics, based on splines (spline2d) and distance
correlation (dcor2d). Give definitions here.. We repeat the study from
before, we first test in Figure ADDREF the behaviour of the index values
over 100 interpolated projections that are randomly selected via the
grand tour algorithm (the same projections as considered above). Note
that here the measures are calculated on non-binned data. The general
behaviour seems similar to the scagnostics monotonic index. The observed
behaviour is smooth (much in contrast to the behaviour of the
scagnostics measures) and we next test the behaviour in a guided tour,
comparing again also the first and final projection. (which index to
select, will try both below\ldots{}) It seems they both maximise
correlation, this might not be too useful?

@@

Physics: what is final projection, what is special about outlying
points?
